---
title: " mapping to a small area"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(fig.width=12, fig.height=8, fig.path=paste0('output',"/"),
                      echo=F, warning=FALSE, message=FALSE, dev = "png", include = T)
```

Let's see the global model making prediction at an area.



packages to install
```{r}
ipak <- function(pkg){

  new.pkg <- pkg[!(pkg %in% installed.packages()[, "Package"])]
  if (length(new.pkg))
    install.packages(new.pkg, dependencies = TRUE, repos='http://cran.muenster.r-project.org')
  sapply(pkg, require, character.only = TRUE)
}
packages <- c( "raster", "dplyr", "devtools", "rgdal","Matrix","xgboost","ranger", "data.table" , "randomForest", "glmnet" ,"rasterVis", "sf", "tmap", "RColorBrewer"   )
ipak(packages)
install_github("mengluchu/APMtools")
library(APMtools)
```
prepare data
```{r}
# prepare data for mapping
data("global_annual")
y_var = "value_mean"
prestring =  "road|nightlight|population|temp|wind|trop|indu|elev"
varstring = paste(prestring,y_var,sep="|")
inde_var = global_annual%>%merge_roads( c(3, 4, 5), keep = F)%>%na.omit() %>%ungroup()%>%dplyr::select(matches(varstring))
sr = stack("dc.grd")
# give the aggregated layers names -- need to match with the dataframe name
 
#writeRaster(srmerged, filename = "/data/lu01/pheonix_1degree_2020.tiff")
#rc2 <- crop(srmerged, extent(srmerged, 2000, 2000, 400, 400))

```

Mapping using xgboost, which handels missing value in predictors both in the modeling and prediction process
```{r}
# nrounds 1650 and eta 0.003 in practice, current setting for fast demonstration, 
xgb_stack(sr=sr, df_var = inde_var, y_var ="value_mean",xgbname = "densexgb.tif",nrounds = 150, eta = 0.05)
levelplot(raster("densexgb.tif"))
```
 
Comparison between prediction patterns using differernt methods. We use the APMtools::predicLA_RF_XGBtiles, which is implemented for three methods, random forest using randomForest package, Lasso using glmnet package, and xgboost using xgboost prackage. 

file names for the output maps
```{r}
xgbname = "xgbdc.tif"
rfname = "rfdc.tif"
laname = "Ladc.tif"
```

```{r}
set.seed(1)
xgb_rounds = 150 # 1650 in practice, 150 for fast demonstration
rf_ntree = 1000 #  
rf_mtry = 34  # optimized from caret
predicLA_RF_XGBtiles(df = inde_var, rasstack = sr, yname = "value_mean", varstring = varstring, xgbname=xgbname, rfname = rfname, laname = laname, ntree = rf_ntree, mtry = rf_mtry,  max_depth = 4, eta = 0.05, nthread = 4, gamma = 1,  nrounds = xgb_rounds) 
```


Visualize
```{r}
myTheme = rasterTheme(region = rev(brewer.pal(7, "Spectral")))# reverse spectral, so that higher values have red color

rfs = raster(rfname)
rfs%>%levelplot(at =seq(20, 50, by =3) , par.settings = myTheme)

xgbs = raster(xgbname)
xgbs%>%levelplot(at = seq(15, 60, by =3) , par.settings = myTheme)

las = raster(laname)
levelplot(las)
stack(rfs, xgbs )%>%levelplot(at = seq(20, 45, by =3), par.settings = myTheme, names.attr = c("random forest","xgboost"))

#png("dc10km.png")
#levelplot(stack(rfs, xgbs))
#dev.off()


```

Visualize using Openstreetmap,show on openStreetMap -importantly, need to reproject to the wgs84 epsg: 4326

```{r}
locations_sf = st_as_sf(global_annual, coords = c("long","lat"))
rfs = projectRaster(rfs, crs=   CRS("+init=epsg:4326"))
xgbs = projectRaster(xgbs, crs=   CRS("+init=epsg:4326"))
las = projectRaster(las, crs=   CRS("+init=epsg:4326"))

osm_valuemean = tm_shape(rfs)+
  tm_raster(names(rfs),palette = "-Spectral", n = 9,alpha = 0.9)+
  tm_shape(xgbs)+tm_raster(names(xgbs),"-Spectral", n = 9,alpha = 0.9)+
  tm_shape(las)+tm_raster(names(las),"-Spectral", n = 9,alpha = 0.9) +
  tm_shape(locations_sf) +
  tm_dots( "value_mean", col = "value_mean", size = 0.05,title = "NO2 value",
           popup.vars = c("value_mean" ))+ 
       tm_view(basemaps = c('OpenStreetMap'))
tmap_save(osm_valuemean, "predictions_pre.html")

```

APMtools:: ranger_stack predicts using a ranger model, as it seems the raster::predict still does not support ranger (actually it explicitly support only randomForest and gbm, others it gives it a try), there are other work-arounds. Here the APMtools::ranger_stack function turns rasterStack into matrix for caluclation and put predictions back to rasters. 
```{r, eval = F}
ranger_pre = ranger_stack(inde_var, "value_mean", sr, 34, 1000)
levelplot(ranger_pre)
```

Compare also ranger and randomForest prediction.
 
```{r, eval = F}


ranger_pre = projectRaster(ranger_pre, crs=   CRS("+init=epsg:4326"))

levelplot(stack(rfs, ranger_pre))
levelplot(rfs- ranger_pre)
values(rfs- ranger_pre)%>%summary() # 0.1 maybe not worth worrying
```

