---
title: "openairquality"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = F)
```

```{r}
ipak <- function(pkg){
 
   new.pkg <- pkg[!(pkg %in% installed.packages()[, "Package"])]
   if (length(new.pkg)) 
    install.packages(new.pkg, dependencies = TRUE, repos = "https://cloud.r-project.org")
  sapply(pkg, require, character.only = TRUE)
}
package = c("ropenaq", "knitr", "xts", "PerformanceAnalytics")
ipak(package)
```
Pollutant variables:
pm25
pm10
no2
so2
o3
co
bc

time step:
hourly
```{r}
sessionInfo()
```
 

```{r}

countries_table <- aq_countries()
kable(countries_table)

```



```{r, eval=F}
cities_table <- aq_cities()
kable(head(cities_table))

```

stations in Netherlands
```{r}
cities_tableIndia <- aq_cities(country="NL", page = 1)
kable(cities_tableIndia)
```

Stations in Utrecht
```{r}
locations_chennai <- aq_locations(country = "NL", city = "Utrecht", parameter = "no2")
kable(locations_chennai)
```


 In October 2017 the API introduced a rate limit of 2,000 requests every 5 minutes. Please keep this in mind. In the case when the request receives a response status of 429 (too many requests), the package will wait 5 minutes.

 
Note, since November 2017 the API only provides access to the last
three months 
Otherdata via Amazon S3
(https://medium.com/@openaq/changes-to-the-openaq-api-and-how-to-access-thefull-archive-of-data-3324b136da8c),
potentially using the aws.s3 package.
 
https://openaq-data.s3.amazonaws.com/index.html
 
There seems to be a problem with the aq_measurements function
```{r, eval=FALSE}
results_table <- aq_measurements(country = "NL", city = "Utrecht", parameter = "no2" , limit = 20, date_from="2020-07-20", page = 1)
kable(results_table)
 
results_table <- aq_measurements(country = "NL", city = "Utrecht", parameter = "no2"  , date_from="2020-05-01", page = 1)
U_NO2 = data.frame(results_table)
table(U_NO2$location)
 
G_NO2= U_NO2[U_NO2$location=="Utrecht-Griftpark", ]
NO2 =  xts(G_NO2$value, G_NO2$dateUTC)
NO2[NO2<0]=NA
#xts::plot.xts(NO2,plot.type="m", at = "chic") 
PerformanceAnalytics::chart.TimeSeries(NO2)
#NO2
 
dit = diff(time(NO2)) 
plot(time(NO2)[-1], dit , main = "missing data in a time series", ylab = "time interval (hour)",xlab = "")
length(diff(time(NO2)))
```

